Collective behavior of Janus Particles Suspended in a Viscous Fluid

### 2022-07-25 doc/PRF2022/PRF_main.tex
started section II

\subsection{Mobility problem}
A suspension of $N_b$ Janus particles is immersed in a viscous solvent.
The domain $\Omega(t) \subset \mathbb{R}^2$ is the solvent phase and $t$
is time. The particles are disks of radius $c$, center $\aa_i$, and
orientation $\theta_i$ relative to the horizontal axis. The boundary of
$\Omega$ is $\bd\Omega = \Gamma_1 \cup \cdots \cup \Gamma_{N_b}$, where
$\Gamma_i$ is the boundary of Janus particle $i$. Assuming inertial
terms are negligible, we have 
\begin{alignat}{3}
\label{eq:stokes}
  -\mu \Delta \uu + \nabla p &= \mathbf{0}, && \xx \in \Omega, \\
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
where $\uu$ is the velocity of the solvent, $p$ is the pressure,
$\uu_\infty$ is the background flow velocity, and $\mu$ is the constant
viscosity. The solvent velocity satisfies the no-slip boundary condition
for a rigid body motion
\begin{equation}
  \label{eq:rigid-vel}
  \uu(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, 
    \quad \xx \in \Gamma_i,
\end{equation}
where $\vv_i$ is the translational velocity, $\omega_i$ is the angular
velocity, and $\langle x, y \rangle^{\perp} = \langle -y, x\rangle$. 
Without inertia, the system in \eqref{eq:stokes}-\eqref{eq:rigid-vel} also satisfies the force-free and torque-free conditions.
This means that the total force $\FF_i,$ and torque $T_i$ from the hydrodynamic stress on each particle (with subscript `$i$')
are in balance with the force and torque from the hydrophobic potential \cite{Fu2022_JFM}.
The total force and torque from the fluid flow are computed as
\begin{equation}
  \label{eq:force}
 \FF_i \equiv \int_{\Gamma_i} \ssigma \cdot \nnu_i \, \dif s, \quad
 T_i \equiv \int_{\Gamma_i} (\xx - \aa_i)^\perp \cdot 
  (\ssigma \cdot \nnu_i) \, \dif s,
  \quad i=1,\ldots,N_b,
\end{equation}
where 
%$\FF_i \in \mathbb{R}^2$ and $T_i \in \mathbb{R}$
%are interparticle forces and torques. 
$\ssigma = -p \mathbf{I} + \mu \left(\nabla \uu + \nabla \uu^T
\right)$ is the hydrodynamic stress tensor (pressure tensor), and
$\nnu_i$ is the unit outward normal on the $i$th particle.

$\FF_i$ and $T_i$ are to be in balance with the force and torque from the hydrophobic interactions between the
particles and the viscous solvent. As in \citet{Fu2022_JFM} such a hydrophobic attraction potential $u(\xx,t)$ is assumed to
the minimizer of the free energy
\begin{equation}
\label{eq:free_energy}
F = \gamma
\int_{\Omega} \rho |\nabla u|^2 + \rho^{-1} u^2
\,d\xx
+ \frac{M}{2}
\sum_{j \neq i} 
P\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right), \quad
\end{equation}
where the hydrophobic attraction potential $u(\xx,t)$ can be regarded as an order parameter for the structure of water,
$\rho$ is a decay length, and $\gamma$ is an interfacial tension. 
$c$ is an ``effective" distance between particle centers, below which the steric repulsion between a pair of JPs becomes important.
In the followings we use the radius of JP for $c$.

The governing equation for $u(\xx,t)$ is obtained by computing the domain variation of the free energy $F$ \cite{Fu20}:
We obtain
a screened Laplace equation boundary value problem as a model for the hydrophobic interaction between water and JPs:
\begin{gather}
  \label{eq:SL}
  -\rho^2 \Delta u + u = 0,\quad \xx \in \Omega, \\
  u = g, \quad \xx \in \bd\Omega, \quad 
  u \rightarrow 0, \quad |\xx| \rightarrow \infty.
\end{gather}
With 
The boundary condition $g$ gives the degree of hydrophobicity the
particle-solvent interface.

Variation of the the free energy \eqref{eq:free_energy} provides
\begin{equation}
  \FF_i = \int_{\Gamma_i} \mathbf{T}\nnu \, \dif s
  - \frac{M}{\rho_0}
  \sum_{j \neq i}
  \frac{\aa_i - \aa_j}{|\aa_i - \aa_j|}
P'\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right)
  ,\quad
T_i = \int_{\Gamma_i} (\xx - \aa_i)^{\perp}\cdot (\mathbf{T}\nnu) \, \dif s, 
\end{equation}
for $i = 1,\dots,N_b$. The term 
$
\mathbf{T}
= \gamma \left[ \rho^{-1} u^2 \mathbf{I}
  + \rho \left(|\nabla u|^2 \mathbf{I} - 2\nabla u \nabla u^T\right)\right]
$
is the hydrophobic stress identified in ~\cite{Fu20} and derived by
variation of the domain techniques.

The dimensionless repulsion profile $P$ takes the form $P(s) = 1 -
\sin(\pi s/2)$ for $0 \leq s < 1$ and $P(s) = 0$ for $1 \leq s <
\infty$.  As such, only those particles within a distance $\rho_0$ of
$\Gamma_i$ contribute to repulsion.  The parameter $M$ is the repulsion
modulus.

We use the parameters $\mu = 1$ cP, $\gamma = 4$ pN nm$^{-1}$,
$\rho = 5$ nm, $c = 1.25$ nm, $M = 2$ pN, and $\rho_0 = 0.5$ nm.

\subsection{Boundary integral represenation and time stepping}
We recast \eqref{eq:stokes}--\eqref{eq:rigid-vel} and \eqref{eq:SL}
as boundary integral equations (BIEs) and discretize each BIE at
$N$ points on each of the $N_b$ particles with a collocation method.
To express the solution of \eqref{eq:SL},
we adopt the double layer potential
\begin{align}
\label{eq:SL_BIE}
u({\bf x}) = \DD[\sigma](\xx) = \int_\Gamma \frac{\partial G(\xx-\yy)}{\partial \nnu_\yy}\sigma(\yy)\, \dif s_\yy
, \quad \xx \in \Omega,
\end{align}
where $G(\xx) = \frac{1}{2\pi}K_0(|\xx|/\rho)$ is the fundamental solution to the screened Laplace equation \eqref{eq:SL},
$K_0$ is the zeroth-order modified Bessel function of the first kind,
and $\sigma$ is a scalar-valued density function.
The subscript in $\dif s_\yy$ denotes integration with respect to $\yy \in \Gamma$. 
For the velocity, we use the representation [REF]
\begin{align}
  \label{eq:velocity}
  \uu(\xx) = \uu_\infty(\xx) + \DD[\eeta](\xx) + 
    \sum_{i=1}^{N_b} \left(\SS(\xx,\aa_i) \cdot \FF_i + 
    \RR(\xx,\aa_i) T_i\right), \quad \xx \in \Omega,
\end{align}
with the vector-valued density function $\eeta$,
Stokeslet 
$\SS(\xx,\aa_i) = \frac{1}{4\pi} \left(-\log |\rr|\mathbf{I} + \frac{\rr \otimes \rr}{|\rr|^2}\right)$,
and Rotlet
$\RR(\xx,\aa_i) = \frac{1}{4\pi} \frac{\rr^\perp}{|\rr|^2}$,
$\rr = \xx - \aa_i$ [REF].

>>>>

\subsection{Mobility problem}
A suspension of $N_b$ Janus particles is immersed in a viscous solvent.
The domain $\Omega(t) \subset \mathbb{R}^2$ is the solvent phase and $t$
is time. The particles are disks of radius $c$, center $\aa_i$, and
orientation $\theta_i$ relative to the horizontal axis. The boundary of
$\Omega$ is $\bd\Omega = \Gamma_1 \cup \cdots \cup \Gamma_{N_b}$, where
$\Gamma_i$ is the boundary of Janus particle $i$. Assuming inertial
terms are negligible, we have 
\begin{alignat}{3}
\label{eq:stokes}
  -\mu \Delta \uu + \nabla p &= \mathbf{0}, && \xx \in \Omega, \\
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
where $\uu$ is the velocity of the solvent, $p$ is the pressure,
$\uu_\infty$ is the background flow velocity, and $\mu$ is the constant
viscosity. The solvent velocity satisfies the no-slip boundary condition
for a rigid body motion
\begin{align}
  \label{eq:rigid-vel}
  \uu(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, 
    \quad \xx \in \Gamma_i,
\end{align}
where $\vv_i$ is the translational velocity, $\omega_i$ is the angular
velocity, and $\langle x, y \rangle^{\perp} = \langle -y, x\rangle$.
Without inertia, the system in~\eqref{eq:stokes}-\eqref{eq:rigid-vel}
also satisfies the force-free and torque-free conditions.
%This means
%that the total force $\FF_i,$ and torque $T_i$ from the hydrodynamic
%stress on each particle (with subscript `$i$') balance the force and
%torque from the hydrophobic potential~\cite{Fu2022_JFM}. 
The total force
and torque from the fluid flow are
\begin{align}
  \label{eq:force}
 \FF_i = \int_{\Gamma_i} \ssigma \cdot \nnu_i \, \dif s, \quad
 T_i = \int_{\Gamma_i} (\xx - \aa_i)^\perp \cdot 
  (\ssigma \cdot \nnu_i) \, \dif s,
  \quad i=1,\ldots,N_b,
\end{align}
respectively, where $\ssigma = -p \mathbf{I} + \mu \left(\nabla \uu +
\nabla \uu^T \right)$ is the hydrodynamic stress tensor (pressure
tensor), and $\nnu_i$ is the unit outward normal on the $i$th particle.

The total force $\FF_i,$ and torque $T_i$ from the hydrodynamic
stress on each particle (with subscript `$i$') balance the force and
torque from the hydrophobic potential. 
%$\FF_i$ and $T_i$ are to be in balance with the force and torque from the hydrophobic interactions between the
%particles and the viscous solvent. 
As in~\citet{Fu2022_JFM} such a hydrophobic attraction potential
$u(\xx,t)$ is assumed to the minimizer of the free energy
\begin{align}
\label{eq:free_energy}
F = \gamma
  \int_{\Omega} \left(\rho |\nabla u|^2 + \rho^{-1} u^2 \right)
\,d\xx
+ \frac{M}{2}
\sum_{j \neq i} 
P\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right), \quad
\end{align}
where the hydrophobic attraction potential $u(\xx,t)$ can be regarded as an order parameter for the structure of water,
$\rho$ is a decay length, and $\gamma$ is an interfacial tension. 
$c$ is an ``effective" distance between particle centers, below which the steric repulsion between a pair of JPs becomes important.
In the followings we let $c$ be the radius of each JP.

The governing equation for $u(\xx,t)$ is obtained by computing the
domain variation of the free energy $F$~\cite{Fu20}. We obtain a
screened Laplace equation boundary value problem as a model for the
hydrophobic interaction between water and JPs:
\begin{alignat}{3}
  \label{eq:SL}
  -\rho^2 \Delta u + u &= 0,\qquad &&\xx \in \Omega, \\
  u &= g, &&\xx \in \bd\Omega, \\
  u &\rightarrow 0, &&|\xx| \rightarrow \infty,
\end{alignat}
where the boundary condition $g$ gives the degree of hydrophobicity the
particle-solvent interface.

The hydrophobic stress is
\begin{align}
\mathbf{T}
= \gamma \left[ \rho^{-1} u^2 \mathbf{I}
  + \rho \left(|\nabla u|^2 \mathbf{I} - 2\nabla u \nabla u^T\right)\right]
\end{align}
and is derived by applying variation of domain techniques to the free
energy~\eqref{eq:free_energy}~\cite{Fu20}. Then, the hydrophobic force
and torque are
\begin{align}
  \FF_i = \int_{\Gamma_i} \mathbf{T}\nnu \, \dif s
  - \frac{M}{\rho_0}
  \sum_{j \neq i}
  \frac{\aa_i - \aa_j}{|\aa_i - \aa_j|}
P'\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right)
  ,\quad
T_i = \int_{\Gamma_i} (\xx - \aa_i)^{\perp}\cdot (\mathbf{T}\nnu) \, \dif s, 
\end{align}
which balance the hydrodynamic force and torque~\eqref{eq:force}, for $i
= 1,\dots,N_b$. 
%The term 
%\begin{align}
%\mathbf{T}
%= \gamma \left[ \rho^{-1} u^2 \mathbf{I}
%  + \rho \left(|\nabla u|^2 \mathbf{I} - 2\nabla u \nabla u^T\right)\right]
%\end{align}
%is the hydrophobic stress~\cite{Fu20} and derived by variation of the
%domain techniques.

Finally, the dimensionless repulsion profile $P$ takes the form $P(s) =
1 - \sin(\pi s/2)$ for $0 \leq s < 1$ and $P(s) = 0$ for $1 \leq s <
\infty$.  As such, only those particles within a distance $\rho_0$ of
$\Gamma_i$ contribute to repulsion. The parameter $M$ is the repulsion
modulus.

We use the same parameters from previous work~\cite{Fu2022_JFM}: $\mu =
1$ cP, $\gamma = 4$ pN nm$^{-1}$, $\rho = 5$ nm, $c = 1.25$ nm, $M = 2$
pN, and $\rho_0 = 0.5$ nm.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary integral represenation and time stepping}
We recast \eqref{eq:stokes}--\eqref{eq:rigid-vel} and \eqref{eq:SL}
as boundary integral equations (BIEs) and discretize each BIE at
$N$ points on each of the $N_b$ particles with a collocation method.
To express the solution of \eqref{eq:SL},
we adopt the double layer potential
\begin{align}
\label{eq:SL_BIE}
u({\bf x}) = \DD[\sigma](\xx) = \int_\Gamma \frac{\partial G(\xx-\yy)}{\partial \nnu_\yy}\sigma(\yy)\, \dif s_\yy
, \quad \xx \in \Omega,
\end{align}
where $G(\xx) = \frac{1}{2\pi}K_0(|\xx|/\rho)$ is the fundamental solution to the screened Laplace equation \eqref{eq:SL},
$K_0$ is the zeroth-order modified Bessel function of the first kind,
and $\sigma$ is a scalar-valued density function.
The subscript in $\dif s_\yy$ denotes integration with respect to $\yy \in \Gamma$. 
For the velocity, we use the representation [REF]
\begin{align}
  \label{eq:velocity}
  \uu(\xx) = \uu_\infty(\xx) + \DD[\eeta](\xx) + 
    \sum_{i=1}^{N_b} \left(\SS(\xx,\aa_i) \cdot \FF_i + 
    \RR(\xx,\aa_i) T_i\right), \quad \xx \in \Omega,
\end{align}
with the vector-valued density function $\eeta$,
Stokeslet 
$\SS(\xx,\aa_i) = \frac{1}{4\pi} \left(-\log |\rr|\mathbf{I} + \frac{\rr \otimes \rr}{|\rr|^2}\right)$,
and Rotlet
$\RR(\xx,\aa_i) = \frac{1}{4\pi} \frac{\rr^\perp}{|\rr|^2}$,
$\rr = \xx - \aa_i$ [REF].
